---
title: "Creating Data Frame of ACIs"
format: html
---

Loading Packages and Mapping functions onto WAV files 
```{r}
library(purrr)
library(soundecology)
library(tidyverse)
library(here)
library(stringr)
paths <- list.files(here("wav_files"), pattern = "\\.WAV$",
                    full.names = TRUE)


wav_files <- map(paths, readWave)
AEI_allWAV<-map(wav_files, acoustic_evenness)
BEI_allWAV<-map(wav_files, bioacoustic_index)
ADI_allWAV<-map(wav_files, acoustic_diversity)
ACI_allWAV<-map(wav_files, acoustic_complexity)
ndsi_allWAV<-map(wav_files, soundecology::ndsi)
```

Acoustic Evenness, Biophony, and Bioacoustic Indices 
```{r}
AEI_df<-as.data.frame(AEI_allWAV)|>
  select(starts_with("aei_left"))|>
  pivot_longer(cols = everything(), names_to = "aei_name",
               values_to = "AEI")
BEI_df<-as.data.frame(BEI_allWAV)|>
  select(starts_with("left_area"))|>
  pivot_longer(cols= everything(), names_to = "bei_name", values_to = "BEI")
  
paths_date <-str_remove(paths, paste0(here(), "/wav_files"))

Biophony_all<-as.data.frame(ndsi_allWAV)|>
  select(starts_with("biophony_left"))|>
  pivot_longer(cols = everything(), names_to = "biop_name", values_to = "Biophony")

```

Full table containing indicies

```{r}

full_WAV_df<-cbind(paths_date, AEI_df, BEI_df, Biophony_all)|>
  separate(col = paths_date, into = c("Date_time", "nonsense"), sep = -4)|>
  select(-(ends_with("name")),-2)

full_WAV_df
```

Obtaining left values for Acoustic Diveristy and Acoustic Complexity indices 
```{r}
ACI_all<-vector("list", length(c(1:10)))
ADI_all<-vector("list", length(c(1:10)))
n<-10
for (i in 1:n){
    ACI_all[i]<-(as.data.frame(ACI_allWAV[[i]]$aci_fl_left_vals))
    ADI_all[i]<-(as.data.frame(ADI_allWAV[[i]]$left_band_values))
} 


ADI_df<-unnest(as.data.frame(ADI_all))
ACI_df<-unnest(as.data.frame(ACI_all))

```

Creating an Acoustic diversity data frame
```{r}
ADI_names<-c("ADI_1", "ADI_2", "ADI_3", "ADI_4", "ADI_5", "ADI_6", "ADI_7", "ADI_8","ADI_9", "ADI_10")
old_adi<-colnames(ADI_df)
ADI_df
ADI_newname<-ADI_df|>rename(cols = old_adi)

ADI_2<-cbind(c(1:10),ADI_2)
ADI_2<-ADI_newname|>rename("ADI_1" = cols1, "ADI_2" = cols2, "ADI_3" = cols3, "ADI_4" = cols4, "ADI_5" = cols5, "ADI_6" = cols6, "ADI_7" = cols7, "ADI_8" = cols8, "ADI_9" = cols9, "ADI_10" = cols10)

ADI_3<-ADI_2|>pivot_longer(cols = c(2:11), names_to = "WAV", values_to = "ADI")|>
  mutate(WAV = as.factor(WAV))|>
  mutate(WAV = fct_relevel(WAV,c("ADI_1", "ADI_2", "ADI_3", "ADI_4", "ADI_5", "ADI_6", "ADI_7", "ADI_8","ADI_9", "ADI_10")))

```

Creating a line plot for Acoustic Diversity 
```{r}
ggplot(data = ADI_3, aes(x=row_number(ADI_3), y = ADI))+
  geom_line(col = "green")+
  facet_wrap(~WAV)+
  theme_minimal()

```

Making Acoustic Complexity data frame 
```{r}
old_aci<-colnames(ACI_df)

ACI_df<-ACI_df|> rename(cols = old_aci)
ACI_df<-ACI_df|>rename("ACI_1" = cols1, "ACI_2" = cols2, "ACI_3" = cols3, "ACI_4" = cols4, "ACI_5" = cols5, "ACI_6" = cols6, "ACI_7" = cols7, "ACI_8" = cols8, "ACI_9" = cols9, "ACI_10" = cols10)

ACI_df<-cbind(Val_ACI,ACI_df)

ACI_df<-ACI_df|>pivot_longer(cols = c(2:11), names_to = "WAV", values_to = "ACI")

ACI_df$WAV<-as.factor(ACI_df$WAV)
ACI_df$WAV<-fct_relevel(ACI_df$WAV, c("ACI_1", "ACI_2", "ACI_3", "ACI_4", "ACI_5", "ACI_6", "ACI_7", "ACI_8","ACI_9", "ACI_10"))
```

Creating a line plot for each WAV File
```{r}
ggplot(data=ACI_df, aes(x=row_number(ACI_df), y=ACI))+
  geom_line(col="orange")+
  theme_minimal()+
  labs(x = "Time(s)", y = "Acoustic Complexity")+
  facet_wrap(~WAV)
```

Acoustic complexity does not seem to match completely with the WAV files, specifically the large spike at the beginning at towards the end of all the graphs. 

```{r}
arrange(ACI_df, ACI)
```

All lowest ACIs fall at "position 176"


starts_with()
Rownumbers()

