---
title: "Analyzing Audio Patterns During the 2024 Total Solar Eclipse"
author: "Kaitlin Heintzman"
institute: "St. Lawrence University"
date: April 25, 2025
format:
  revealjs:
    transition: fade
    theme: serif 
    slide-number: true
    footer: "Total Solar Eclipse Sound Analysis"
embed-resources: true
---
```{r}
library(tidyverse)
library(lubridate)
library(hms)
library(forecast)
library(soundecology)
library(here)
library(mgcv)
library(modelr)
library(pander)
library(knitr)
```

```{r}
A001_SD001<-readRDS(here("rds_files/A001_SD001.rds"))
A002_SD013<-readRDS(here("rds_files/A002_SD013.rds"))
A003_SD005<-readRDS(here("rds_files/A003_SD005.rds"))
A004_SD012<-readRDS(here("rds_files/A004_SD012.rds"))
A005_SD002<-readRDS(here("rds_files/A005_SD002.rds"))
A006_SD006<-readRDS(here("rds_files/A006_SD006.rds"))
A007_SD017<-readRDS(here("rds_files/A007_SD017.rds"))
A008_SD007<-readRDS(here("rds_files/A008_SD007.rds"))
A009_SD009<-readRDS(here("rds_files/A009_SD009.rds"))
A010_SD014<-readRDS(here("rds_files/A010_SD014.rds"))
A011_SD018<-readRDS(here("rds_files/A011_SD018.rds"))
A013_SD016<-readRDS(here("rds_files/A013_SD016.rds"))
A014_SD021<-readRDS(here("rds_files/A014_SD021.rds"))
A015_SD010<-readRDS(here("rds_files/A015_SD010.rds"))
A016_SD022<-readRDS(here("rds_files/A016_SD022.rds"))
A017_SD024<-readRDS(here("rds_files/A017_SD024.rds"))
A018_SD011<-readRDS(here("rds_files/A018_SD011.rds"))
A019_SD008<-readRDS(here("rds_files/A019_SD008.rds"))
A021_SD023<-readRDS(here("rds_files/A021_SD023.rds"))
A022_SD019<-readRDS(here("rds_files/A022_SD019.rds"))

eclipse_start<-hms(00,00,14)
eclipse_end<-hms(00,50,16)
```


## Outline

* Introduction
  
* Data Collection 
  
* Data Frame Creation 
  
* GAM
  
* Conclusion 

## Introduction 

* Our goal was to look deeper into effects that the total solar eclipse has on wildlife activity. To do so, we used generalized additive models to observe 5 indices across the time of the eclipse and we found temporal patterns in 3 of these indices, suggesting that there is audiological variation during the eclipse.

* This was an amazing and unique opportunity to look into such a rare phenomenon and to use data that was collected by SLU faculty. 
	
## Audio Recorder Placement

	(map)
	
	20 audio recorders   
	
	ArcGIS was used to find forested wetland habitats so both avian and amphibian wildlife would be sampled
	

Heavy inspiration for this analysis was taken from: 

Gerber, J. E., Howard, D., & Quinn, J. E. (2020). Soundscape shifts during the 2017 total solar eclipse: An application of dispersed automated recording units to study ephemeral acoustic events. Biodiversity, 21(1), 41–47. https://doi.org/10.1080/14888386.2020.1715834 
## Data Collection {.smaller}
	  
* Recorders collected audio from March 30th to April 16th

| Times         | Reasoning                                        |
|---------------|--------------------------------------------------|
| 5:45 - 7:15   | 30-45 minutes before sunrise                     |
| 14:00 - 16:00 | full period of time corresponding to the eclipse |
| 19:00 - 20:00 | \~30 minutes before and after sunset             |
| 23:00 - 23:30 | sample nocturnal activity                        |

: {.striped}


<br>

* Recorders sampled 55 seconds on and 5 seconds off

* This produced a total of 115,610 audio clips
	  
## Audio Indices {.smaller}

* <span style="background-color: lightgreen; color: black; padding: 2px 4px; border-radius: 3px;">Bioacoustic Index </span> -  Assesses diversity and abundance of specifically biological noise in relation to the period of the file with the lowest sound. 

* <span style="background-color: coral; color: black; padding: 2px 4px; border-radius: 3px;"> Biophony </span>  - Measures the average frequency of biological sound

* Acoustic Complexity - Focused on expressing the spatial and temporal complexity in sound by capturing the intensity and frequency shifts across an audio file. 

* Acoustic Diversity - Generates proportions over set intervals to assess evenness and variety in sound distribution 

* Acoustic Evenness - Measure how sound is distributed across frequency bands, assess equality and inequality of the rate that sound is produced from a source

<br>

Villanueva-Rivera LJ, Pijanowski BC (2018). _soundecology: Soundscape Ecology_. R package
  version 1.3.3, <https://CRAN.R-project.org/package=soundecology>.

## Visualize and Youtube 

<https://youtu.be/3dftvaUn--Q> 
	
```{r}
#| echo: false
eclipse_time<-A001_SD001|> filter(hour>= eclipse_start & hour<= eclipse_end)
eclipseonly<-A001_SD001|> filter(hour>= eclipse_start & hour<= eclipse_end)|> filter(day == "2024-04-08")

single_day<-A001_SD001|>filter(day == "2024-04-09")

dawn<- single_day|> filter(hour >= as_hms("05:45:00") & hour <= as_hms("07:15:00"))
eclipse<- single_day|> filter(hour >= as_hms("14:00:00") & hour <= as_hms("16:00:00"))
afternoon<- single_day|> filter(hour >= as_hms("19:00:00") & hour <= as_hms("20:00:00"))
night<- single_day|> filter(hour >= as_hms("23:00:00") & hour <= as_hms("23:30:00"))

ggplot(data = dawn, aes( x = hour, y = bei))+
  geom_rect( aes(xmin = as_hms("05:45:00"), xmax = as_hms("07:15:00"), ymin = 0, ymax = 4), fill = "lightblue1")+
  geom_rect(data = dawn, aes(xmin = as_hms("14:00:00"), xmax = as_hms("16:00:00"), ymin = 0, ymax = 4), fill = "lightblue1")+
  geom_rect(data = dawn, aes(xmin = as_hms("19:00:00"), xmax = as_hms("20:00:00"), ymin = 0, ymax = 4), fill = "lightblue1")+
  geom_rect(data = dawn, aes(xmin = as_hms("23:00:00"), xmax = as_hms("23:30:00"), ymin = 0, ymax = 4), fill = "lightblue1")+
  geom_line(col = "darkblue")+
  geom_line(data = eclipse, aes(x = hour, y = bei), col = "darkblue")+
  geom_line(data = afternoon, aes(x = hour, y = bei), col = "darkblue")+
  geom_line(data = night, aes(x = hour, y = bei), col = "darkblue")+
  theme_minimal()+
  facet_wrap(~day)+
  labs( title = "Example of Full Pattern - Audio Recorder 1")+
  scale_x_continuous(name = "Hour", breaks = c(hms(00,45,5), hms(00,00,14), hms(00,00, 19),hms(00,00,23)))+
  theme(axis.text.x = element_text(angle = 45, vjust= 0.5), axis.text = element_text(size = 20))
  
```

## Using R and HPC {.smaller}


* Function to take sound files and return cleaned data frame with information needed for analysis 
	 
```{r}
table<- A001_SD001|> mutate(ACI = "<dbl [256]>", ADI = "<dbl [10]>")|>
  select(-ACI_all,-ADI_all)
head(table)|>kable(format = "html")|>
  kableExtra::kable_styling(font_size = 18)|>
  kableExtra::kable_classic_2()
```

<br>

* Over a **terabyte** of data

* Due to massive size, we used St. Lawrence's HPC to run and store our files incrementally. 

* With the HPC, each folder took 10 and a half hours to process. 
	
## Generalized Additive Modeling {.smaller}
  
$$ Index \sim spline(day, by = folder) + spline(folder, bs = "re") + day$$

* Fit a smoothing function for day, and have it vary by folder

* Adjust for folder differences, but treat this as a random effect

* Incorporate a linear trend for day

```{r}
#| echo: false
fullAudio<-rbind(A001_SD001, A002_SD013, A003_SD005, A004_SD012,A005_SD002, A006_SD006, A007_SD017, 
                 A008_SD007, A009_SD009, A010_SD014, A011_SD018, A013_SD016,A014_SD021, A015_SD010, A016_SD022,
                 A017_SD024, A018_SD011, A019_SD008, A021_SD023, A022_SD019)|>
  group_by(folder_name)|>
  mutate(fullACI = sapply(ACI_all,sum))|>
  mutate(fullADI = sapply(ADI_all, sum))


modelEclipse_df<-fullAudio|> mutate(eclipse = ifelse(day == "2024-04-08", 
                                                     "eclipse", 
                                                     "not_eclipse"))|>
  mutate(eclipse = as_factor(eclipse), folder_name = as_factor(folder_name))|>
  mutate(hour_numeric = as.numeric(hour))|>
  filter(hour>= eclipse_start & hour<= eclipse_end &
           day %in% c("2024-04-06", "2024-04-07", "2024-04-08",
                    "2024-04-09", "2024-04-11"))|>
  mutate(day_factor = as.factor(day))|>
  select(hour, hour_numeric, everything())
bei_mod<- mgcv::gam(bei~ s(hour_numeric, by = day_factor)+
                      day_factor + s(folder_name, bs = "re"),
                    data = modelEclipse_df)

```
	
::: {.callout-tip} 

## Why GAM?
  We can maintain additivity while incorporating non-linear functions!
:::

* If time, EDF note? 

## Bioacoustic Model Visualization {.smaller}

```{r}
#| echo: false
#| fig-align: center 

partial_begin = tibble(bei = (3.5), biophony = (2.5), fullACI = (1950), 
                       fullADI = (9), aei = (0.7), hour_numeric = hms(38,11,14))
second_df = tibble(bei = (3.5), biophony = (2.5), fullACI = (1950),
                   fullADI = (9), aei = (0.7), hour_numeric = hms(00,45,14))
totality = tibble(bei = (3.5), biophony = (2.5), fullACI = (1950),
                  fullADI = (9), aei = (0.7),  hour_numeric = hms(52,23,15))
fourth_df = tibble(bei = (3.5), biophony = (2.5), fullACI = (1950), 
                   fullADI = (9), aei = (0.7), hour_numeric = hms(00,00,16))
partial_end = tibble(bei = (3.5), biophony = (2.5), fullACI = (1950),
                     fullADI = (9), aei = (0.7), hour_numeric = hms(38,35,16))

grid <- modelEclipse_df |> 
  ungroup() |>
  data_grid(hour_numeric = seq_range(hour_numeric, n = 40), folder_name = "/A014_SD021",
    day_factor = modelEclipse_df |> pull(day_factor) |> levels(),
  )

gam_aug <- broom::augment(bei_mod, newdata = grid)


# visual across the time of eclipse
ggplot(data = gam_aug, aes(x = hour_numeric, y = .fitted)) +
  geom_point(data = modelEclipse_df, aes(y = bei, colour = day_factor), alpha = 0.1) +
  geom_line(aes(colour = day_factor), linewidth = 2) +
  geom_point(data = totality, aes(x = hour_numeric, y = bei), pch = 15, size = 6)+
  geom_point(data = partial_begin, aes(x = hour_numeric, y = bei), pch = 15, size = 6)+
  geom_point(data = partial_begin, aes(x = hour_numeric, y = bei), size = 4, col = "white")+
  geom_point(x = hms(00,10,14), y = 3.5, size = 2)+
  geom_point(data = second_df, aes(x = hour_numeric, y = bei), pch = 15, size = 6)+
  geom_point(data = second_df, aes(x = hour_numeric, y = bei), size = 4, col = "white")+
  geom_point(x = hms(00,44,14), y = 3.5, size = 3)+
  geom_point(data = fourth_df, aes(x = hour_numeric, y = bei), pch = 15, size = 6)+
  geom_point(data = fourth_df, aes(x = hour_numeric, y = bei), size = 4, col = "white")+
  geom_point(x = hms(00,01,16), y = 3.5, size = 3)+
  geom_point(data = partial_end, aes(x = hour_numeric, y = bei),pch = 15 , size = 6)+
  geom_point(data = partial_end, aes(x = hour_numeric, y = bei), size = 4, col = "white")+
  geom_point(x = hms(00,37,16), y = 3.5, size = 2)+
  scale_colour_viridis_d() +
  theme_minimal(base_size = 30) +
  labs(colour = "eclipse_or_not",
       y = "Bioacoustic Index")+
  scale_x_continuous(name = "Time", breaks = c(eclipse_start, hms(00,00,15), hms(00,00,16), eclipse_end))

```



## Biophony Model Visualization {.smaller}

```{r}
#| echo: false
biophony_mod<- mgcv::gam(biophony~ s(hour_numeric, by = day_factor)+
                      day_factor + s(folder_name, bs = "re")
                    , data = modelEclipse_df)
grid <- modelEclipse_df |> ungroup() |>
  data_grid(hour_numeric = seq_range(hour_numeric, n = 40), folder_name = "/A001_SD001",
            day_factor = modelEclipse_df |> pull(day_factor) |> levels()
  )

gam_aug <- broom::augment(biophony_mod, newdata = grid)

ggplot(data = gam_aug, aes(x = hour_numeric, y = .fitted)) +
  geom_point(data = modelEclipse_df, aes(y = biophony, colour = day_factor), alpha = 0.1) +
  geom_line(aes(colour = day_factor),
            linewidth = 2) +
  geom_point(data = totality, aes(x = hour_numeric, y = biophony), pch = 15, size = 6)+
  geom_point(data = partial_begin, aes(x = hour_numeric, y = biophony), pch = 15, size = 6)+
  geom_point(data = partial_begin, aes(x = hour_numeric, y = biophony), size = 4, col = "white")+
  geom_point(x = hms(00,10,14), y = 2.5, size = 2)+
  geom_point(data = second_df, aes(x = hour_numeric, y = biophony), pch = 15, size = 6)+
  geom_point(data = second_df, aes(x = hour_numeric, y = biophony), size = 4, col = "white")+
  geom_point(x = hms(00,44,14), y = 2.5, size = 3)+
  geom_point(data = fourth_df, aes(x = hour_numeric, y = biophony), pch = 15, size = 6)+
  geom_point(data = fourth_df, aes(x = hour_numeric, y = biophony), size = 4, col = "white")+
  geom_point(x = hms(00,01,16), y = 2.5, size = 3)+
  geom_point(data = partial_end, aes(x = hour_numeric, y = biophony),pch = 15 , size = 6)+
  geom_point(data = partial_end, aes(x = hour_numeric, y = biophony), size = 4, col = "white")+
  geom_point(x = hms(00,37,16), y = 2.5, size = 2)+
  scale_colour_viridis_d() +
  theme_minimal(base_size = 30) +
  labs(colour = "eclipse_or_not",
       y = "Biophony")+
    scale_x_continuous(name = "Time", breaks = c(eclipse_start, hms(00,00,15), hms(00,00,16), eclipse_end))
```


## Findings {.smaller}

* Based on our analysis we observed potential patterns in 3 of our indices:

  + Bioacoustic Index

  + Acoustic Evenness 

  + Acoustic Diversity 
  
<br>

* These indices showed increases or decreases around the time of the eclipse, specifically with variation around the time of totality. 

* This suggests that wildlife is affected by the sudden changes in light that occurred during the solar eclipse or by some other eclipse-driven change. 


## Future Projects

* Accounting for temporal autocorrelation: (Emilia Agostinelli '26)

  + Incorporating time into the model 
  
  + Utilizing time series approaches  
  
<br>

* What species are being affected:  (Aidan Fauth '26)

  + Identifying the wildlife present in the clips
  
  + Machine learning possibility 
	
## Acknowledgements 

* Thank you to Erika Barthelmess and her team of students who made this project possible by collecting the data. 
  
* Thank you to the MCSS department and specifically Matt Higham for guiding me through this project.
  
* Special Thanks to Ed Harcourt and Lisa Torrey for assistance in using and understanding the HPC process. 


## to do: 

add map



