---
title: "Graphing First WAV and Obtaining Indices"
format: html
---
```{r}
library(soundecology)
library(tuneR)
library(here)
library(ggplot2)
library(readr)
```

### Acoustic Complexity Plot of WAV - 240401_0616100
```{r}
WAV_0616100_240401<-readWave(here("wav_files/20240401_061600.WAV"))

ACI_01<-acoustic_complexity(WAV_0616100_240401)
ACI_01_val<-ACI_01$aci_fl_left_vals
df_ACI<-as.data.frame(ACI_01$aci_fl_left_vals)

plot(x=1:256, y=ACI_01_val, col="orange", pch=16)

ggplot(data=df_ACI, aes(x=1:256, y=ACI_01$aci_fl_left_vals))+
  geom_line(col="orange")+
  theme_minimal()+
  labs(x = "Time(s)", y = "Acoustic Complexity")

```
Not sure about x-axis, how to label for the actual measurement of time present

### Other Indices for this file: 

Bioacoustic Index for WAV - 240401_0616100
```{r}
Bioindex1<-bioacoustic_index(WAV_0616100_240401)
```

Acoustic Diversity Index for WAV - 240401_0616100
```{r}
ADI1<-acoustic_diversity(WAV_0616100_240401)

ADI1_lv<-as.data.frame(ADI1$left_band_values)

ggplot(data=ADI1_lv, aes(x=1:10, y=ADI1$left_band_values))+
  geom_point()+
  theme_minimal()
```

Acoustic Evenness Index for WAV - 240401_0616100
```{r}
AEI_1<-acoustic_evenness(WAV_0616100_240401)
```

Attempting to use the multiple sounds function (on 5 WAV files)

```{r}
multiple_sounds(directory = "wav_files", resultfile = "WAV5_ACI.csv", soundindex = "acoustic_complexity", no_cores="max")
```

```{r}
WAV5_ACI <- read_csv("WAV5_ACI.csv")
```

Returns singular indices but does not return the list of values used to create the acoustic complexity plot like above. "aci_fl_left_vals"

### Making Acoustic Complexity plots for more WAV-Files:

```{r}
WAV_063200_240401<-readWave(here("wav_files/20240401_063200.WAV"))

ACI_02<-acoustic_complexity(WAV_063200_240401)
ACI_02_val<-ACI_02$aci_fl_left_vals
df_ACI2<-as.data.frame(ACI_02$aci_fl_left_vals)

ggplot(data=df_ACI2, aes(x=1:256, y=ACI_02$aci_fl_left_vals))+
  geom_line(col="purple")+
  theme_minimal()+
  labs(x = "Time(s)", y = "Acoustic Complexity")
```

```{r}
WAV_054600_240403<-readWave(here("wav_files/20240403_054600.WAV"))

ACI_03<-acoustic_complexity(WAV_054600_240403)
ACI_03_val<-ACI_03$aci_fl_left_vals
df_ACI3<-as.data.frame(ACI_03$aci_fl_left_vals)

ggplot(data=df_ACI3, aes(x=1:256, y=ACI_03$aci_fl_left_vals))+
  geom_line(col="green")+
  theme_minimal()+
  labs(x = "Time(s)", y = "Acoustic Complexity")
```



